---
title: "MR-MOSS Quick Start"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{MR-MOSS Quick Start}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## Overview

MR-MOSS fits a likelihood-based multi-outcome Mendelian randomization model from
harmonized GWAS summary statistics and a working outcome correlation matrix. The
same fitted object can be used for outcome-specific, subset/domain-level, and
all-outcome likelihood-ratio tests.

This vignette uses only the toy files bundled with the package.

## Load Toy Inputs

```{r}
library(MRMOSS)

summary_path <- system.file("extdata", "toy_summary_stats.csv", package = "MRMOSS")
cor_path <- system.file("extdata", "toy_outcome_correlation.csv", package = "MRMOSS")
domain_path <- system.file("extdata", "toy_domain_map.csv", package = "MRMOSS")

toy <- read_mrmoss_summary_stats(summary_path)
R <- read_outcome_correlation(cor_path)
domain_map <- read.csv(domain_path)

str(toy[c("n1", "n2", "outcomes")])
R
domain_map
```

## Check Inputs

```{r}
checked <- check_mrmoss_inputs(
  gamma_hat = toy$gamma_hat,
  Gamma_hat = toy$Gamma_hat,
  R = R,
  n1 = toy$n1,
  n2 = toy$n2
)

checked$min_eigenvalue
checked$positive_definite
```

## Fit MR-MOSS

```{r}
fit <- fit_mrmoss(summary_stats = toy, R = R, maxiter = 500)

fit$beta
fit$iteration
```

## Likelihood-Ratio Tests

```{r}
outcome_lrt(fit)
global_lrt(fit)
domain_lrt(fit, domain_map)
```

Custom outcome subsets can be tested directly.

```{r}
subset_lrt(
  fit,
  subsets = list(first_two = c("outcome_1", "outcome_2"), third = "outcome_3")
)
```

## Estimate a Working Correlation Matrix

If null-variant outcome z-scores are available, estimate a working correlation
matrix before fitting.

```{r}
set.seed(1)
null_z <- matrix(rnorm(300), ncol = 3)
colnames(null_z) <- toy$outcomes

estimated_R <- estimate_outcome_correlation(null_z)
estimated_R
attr(estimated_R, "nearPD_used")
```

## Next Steps for Real Data

For real data, replace the toy inputs with harmonized SNP-exposure and
SNP-outcome summary statistics, a matching outcome correlation matrix, and an
optional domain map. Confirm that the same outcome names appear in every file.

MR-MOSS expects the real-data summary file to be prepared before fitting:
instrument SNPs should be selected and LD-clumped, all exposure and outcome
betas should be oriented to the row-level effect allele, unresolved palindromic
or mismatched variants should be removed, and an effective scalar exposure and
outcome sample size strategy should be documented. See the README and
`scripts/04_template_real_gwas_analysis.R` for a copy-and-edit real-data
analysis template.
